Real-space computation of $E$/$B$-mode maps I: Formalism, Compact Kernels, and Polarized Filaments

### 2018-07-13 articles/formalism/sections/generalizedeb.tex
Edited moving average, dealt with problematic equation reference

%--------------------------------------------------------
\paragraph{Relation to the spin raising $\eth^2$ and lowering $\bar{\eth}^2$ operators.}
Recall that on operating twice with the spin lowering operator on the Stokes charge ${}_{+2}X$ results in filtered version of $E/B$ maps as in \eq{eq:ebdef}. Now note that it is possible to construct a modified real space operator by choosing the harmonic space function to be $g_{\ell} = ({(\ell+2)!/(\ell-2)!})^{1/2}$, resulting in similarly filtered $E/B$ maps as follows:
%
\beq 
[\mathcal{E} + i \mathcal{B}](\hat{n}_e)=- \Delta \Omega\sum_{q=1}^{N_{\rm pix}} \Bigg\lbrace  \left[  \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} P_{\ell}^2(\beta_{qe}) \right] e^{-i2\alpha_{eq}} {}_{+2}X(\hat{n}_{q}) \Bigg\rbrace \,.\label{eq:bl_ebdef_lower} 
\eeq
%
Comparing \eq{eq:ebdef_lower} to \eq{eq:bl_ebdef_lower} makes apparent the following mapping:
%
\beq
\bar{\eth}^2 \equiv \Delta \Omega \sum_{q=1}^{N_{\rm pix}} \left[ \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} P_{\ell}^2(\beta_{qe}) \right] e^{-i2\alpha_{eq}}\,.
\eeq
%
The operator $\bar{\eth}^2$ is composed of derivative operations and hence has an implicit positional dependence for all numerical purposes which is made explicit in the band limited version of the operator.  The band limited version of the spin raising operator $\eth^2$ is derived by simply taking the conjugate of the above equation.
%--------------------------------------------------------

>>>>

%--------------------------------------------------------
\paragraph{Relation to the spin raising $\eth^2$ and lowering $\bar{\eth}^2$ operators.}
Recall that on operating twice with the spin lowering operator on the Stokes charge ${}_{+2}X$ results in filtered version of $E/B$ maps as in \eq{eq:ebdef}. Now note that it is possible to construct a modified real space operator by choosing the harmonic space function to be $g_{\ell} = ({(\ell+2)!/(\ell-2)!})^{1/2}$, resulting in similarly filtered $E/B$ maps as follows:
%
\beq 
[\mathcal{E} + i \mathcal{B}](\hat{n}_e)=- \Delta \Omega\sum_{q=1}^{N_{\rm pix}} \Bigg\lbrace  \left[  \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} P_{\ell}^2(\beta_{qe}) \right] e^{-i2\alpha_{eq}} {}_{+2}X(\hat{n}_{q}) \Bigg\rbrace \,. \label{eq:bl_ebdef_lower} 
\eeq
%
Comparing to \eq{eq:ebdef_lower} %to \eq{eq:bl_ebdef_lower}
makes apparent the following mapping:
%
\beq
\bar{\eth}^2 \equiv \Delta \Omega \sum_{q=1}^{N_{\rm pix}} \left[ \sum_{\ell=\ell_{\rm min}}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} P_{\ell}^2(\beta_{qe}) \right] e^{-i2\alpha_{eq}}\,.
\eeq
%
The operator $\bar{\eth}^2$ is composed of derivative operations and hence has an implicit positional dependence for all numerical purposes which is made explicit in the band limited version of the operator.  The band limited version of the spin raising operator $\eth^2$ is derived by simply taking the conjugate of the above equation.
%--------------------------------------------------------

### 2018-07-13 articles/formalism/sections/kernel_discussion.tex
Edited moving average, dealt with problematic equation reference

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Quantifying the non-locality of E \& B modes} \label{sec:radial_locality}
\subsection{The non-locality of the real space operators} \label{sec:radial_locality}
%
\begin{figure}[t]
\centering
\includegraphics[width=0.8\columnwidth]{beta_kernel.pdf}
\caption{The figure depicts the radial part of the convolution kernels. These radial function have been evaluated with the band limit fixed at $\ell \in [2,24]$. The vertical dashed line marks the approximate Healpix pixel size of a NSIDE=8, which is the lowest resolution that allows access to $\ell_{\rm max}=24$.}
\label{fig:beta_kernel}
\end{figure}
%
Above, we have explored in detail the  azimuthal dependence of the real space kernels.  Here we probe the radial dependence, which both  determines the non-locality of the operators and encodes all their multipole dependencies. For illustration, \fig{fig:beta_kernel} shows the radial kernels ${_{\mm}f}, {_{\md}f},{_{\mi}f}$, evaluated using the respective multipole sums given in \eq{eq:rad_ker_queb} and \eq{eq:f2_rad_ker} in the band limit $\ell \in [2,24]$. We choose a low band limit to evaluate these radial kernels to highlight some key features of their radial profile.
%
\begin{figure}[t]
\subfigure[\label{fig:fbeta}]{\includegraphics[width=0.325\columnwidth]{rad_ker_fn_of_ellmax.pdf}}\hfill
\subfigure[\label{fig:dfbeta}]{\includegraphics[width=0.325\columnwidth]{rad_ker_d_fn_of_ellmax.pdf}}\hfill
\subfigure[\label{fig:ifbeta}]{\includegraphics[width=0.325\columnwidth]{rad_ker_i_fn_of_ellmax.pdf}}\hfill
\centering
\subfigure[\label{fig:fbeta_avg}]{\includegraphics[width=0.325\columnwidth]{rad_ker_fn_of_ellmax_zaldariagga_scaling.pdf}}
\subfigure[\label{fig:dfbeta_avg}]{\includegraphics[width=0.325\columnwidth]{rad_ker_i_fn_of_ellmax_zaldariagga_scaling.pdf}}

@@

The resultant set of radial function are depicted in \fig{fig:rad_ker_decay}. While the amplitude of these radial function scales up as $\propto \ell_{\rm max}^2$, for clarity in the plot their global maxima are normalized to unity.  This normalization highlights the key feature, that on increasing $\ell_{\rm max}$ the radial kernels shift left, attaining their global maxima at progressively small angular distances $\beta$.  At intermediate values of $\beta$, the envelope of the radial functions is fit well by a power law $ \propto \beta^{-n}$ as seen in \fig{fig:rad_ker_decay}.
In fact these finding are neatly summarized in the observation that the radial functions computed by evaluating the multipole sums to different maximum multipoles are approximately self-similar and follow this telescoping and scaling property: $${}_rf(\beta,2,\ell_{\rm max}) \approx \Big[\frac{\ell_{\rm max}}{\ell'_{\rm max}}\Big]^2{}_rf(\beta'=\frac{\ell'_{\rm max}}{\ell_{\rm max}} \beta ,2,\ell'_{\rm max}) \,,$$ where ${}_rf , r \in [\mm, \md,\mi]$ denotes all the different radial functions. We can now understand the amplitude scaling and the shifting left of the radial kernels on increasing the maximum multipole by studying this telescoping property. Specifically the global maxima of the kernel with maximum multipole $\ell_{max}$ is amplified/supressed by a factor $(\ell_{\rm max}/\ell'_{\rm max})^2$ and the oscillating part is compressed/expanded by a factor $(\ell'_{\rm max}/\ell_{\rm max})^{-1}$ with respect to the kernel defined by the maximum multipole $\ell'_{\rm max}$.

@@

\revisit{In the flat sky limit it has been argued that form of the radial function has to be ${}_\mm f(\beta)\simeq\beta^{-2}$ so as to ensures that the Fourier modes for Stokes parameters $Q/U$ are related  to those of $E/B$ modes merely by rotations and no scale dependent factor \cite{Zaldarriaga2001a}. Here we numerically studied the formally derived radially function ${}_\mm f(\beta)$ which reveals an oscillatory nature with amplitude decaying on increasing angular distance. While we find that the positive envelope of ${}_{\mm}f$ scales as $\beta^{-1.55}$, the function resulting from performing a sliding average on it is seen to scale as $\beta^{-2}$ at intermediate angular separations as seen in \fig{fig:fbeta_avg}, in agreement with the hypothesis made in \cite{Zaldarriaga2001a}. We find that the function resulting from performing a sliding average on ${}_{\md}f$ also shows a similar scaling of $\beta^{-2}$ as seen in \fig{fig:dfbeta_avg}.}

\revisit{In essence the radial function resulting from taking a sliding average can be thought of as transitioning to some lower band limit (determined by the width of the sliding window) radial function from a higher band limit radial function.  On carrying out this moving average on radial functions evaluated with a progressively increasing band limit, one seen that a larger fraction of the radial function matches the scaling of $\beta^{-2}$. Extrapolating this trend to the case of infinite band limit one can see that the behaviour will approach $\beta^{-2}$ across the domain $\beta\in[0,\pi]$.}

\revisit{Near the edges of the domain $\beta \rightarrow 0,\pi$ the window function gets modified and is partly responsible for the scaling deviating from the power law. More importantly note that ${}_{\mm}f$ and ${}_{\md}f$ vanish at $\beta=0$ and therefore the sliding average in the vicinity of this point cannot have a power law behaviour of the form $\beta^{-2}$.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Quantifying the non-locality of E \& B modes} \label{sec:radial_locality}
\subsection{The non-locality of the real space operators} \label{sec:radial_locality}
%
\begin{figure}[t]
\centering
\includegraphics[width=0.8\columnwidth]{beta_kernel.pdf}
\caption{The figure depicts the radial part of the convolution kernels. These radial function have been evaluated with the band limit fixed at $\ell \in [2,24]$. The vertical dashed line marks the approximate Healpix pixel size of a NSIDE=8, which is the lowest resolution that allows access to $\ell_{\rm max}=24$.}
\label{fig:beta_kernel}
\end{figure}
%
Above, we have explored in detail the  azimuthal dependence of the real space kernels.  Here we probe the radial dependence, which both  determines the non-locality of the operators and encodes all their multipole dependencies. For illustration, \fig{fig:beta_kernel} shows the radial kernels ${_{\mm}f}, {_{\md}f},{_{\mi}f}$, evaluated using the respective multipole sums given in \eq{eq:rad_ker_queb} and \eq{eq:f2_rad_ker} in the band limit $\ell \in [2,24]$. We choose such a low band limit to highlight some key features of their radial profile.
%
\begin{figure}[t]
\subfigure[\label{fig:fbeta}]{\includegraphics[width=0.325\columnwidth]{rad_ker_fn_of_ellmax.pdf}}\hfill
\subfigure[\label{fig:dfbeta}]{\includegraphics[width=0.325\columnwidth]{rad_ker_d_fn_of_ellmax.pdf}}\hfill
\subfigure[\label{fig:ifbeta}]{\includegraphics[width=0.325\columnwidth]{rad_ker_i_fn_of_ellmax.pdf}}\hfill
\centering
\subfigure[\label{fig:fbeta_avg}]{\includegraphics[width=0.325\columnwidth]{rad_ker_fn_of_ellmax_zaldariagga_scaling.pdf}}
\subfigure[\label{fig:dfbeta_avg}]{\includegraphics[width=0.325\columnwidth]{rad_ker_i_fn_of_ellmax_zaldariagga_scaling.pdf}}

@@

The resultant set of radial function are depicted in \fig{fig:rad_ker_decay}. While the amplitude of these radial function scales up as $\propto \ell_{\rm max}^2$, for clarity in the plot their global maxima are normalized to unity.  This normalization highlights the key feature, that on increasing $\ell_{\rm max}$ the radial kernels shift left, attaining their global maxima at progressively small angular distances $\beta$.  At intermediate values of $\beta$, the envelope of the radial functions is fit well by a power law $ \propto \beta^{-n}$.  As the band limit changes, the small angle portion of the radial functions have a similar shape.

These finding are neatly summarized in the observation that the radial functions with different maximum multipoles are approximately self-similar and related by this telescoping and scaling property:
$${}_rf(\beta,2,\ell_{\rm max}) \approx \Big[\frac{\ell_{\rm max}}{\ell'_{\rm max}}\Big]^2{}_rf(\beta'=\frac{\ell'_{\rm max}}{\ell_{\rm max}} \beta ,2,\ell'_{\rm max}) \qquad \mbox{(small angles)},$$
where ${}_rf , r \in [\mm, \md,\mi]$ denotes all the different radial functions. This telescoping property encapsulates both the amplitude scaling and leftward shift of the radial kernels on increasing the maximum multipole. Specifically the global maxima of the kernel with maximum multipole $\ell_{max}$ is amplified/supressed by a factor $(\ell_{\rm max}/\ell'_{\rm max})^2$ and the oscillating part is compressed/expanded by a factor $(\ell'_{\rm max}/\ell_{\rm max})^{-1}$ with respect to the kernel defined by the maximum multipole $\ell'_{\rm max}$.

@@

\revisit{The radial function ${}_\mm f(\beta)$ is oscillatory, regardless of the maximum $l_{\rm max}$. The envelope of the oscillation decays with angular separation at intermediate $\beta$ but increases as $\beta \rightarrow \pi$.  We find that the positive envelope of ${}_{\mm}f$ scales as $\beta^{-1.55}$ at intermediate $\beta$.  However, when we average over several oscillations with a moving window, we find that the average behavior as intermediate $\beta$ scales as $\beta^{-2}$, visible in \fig{fig:fbeta_avg}.  (We have narrowed the window as the band limit increases so that the number of oscillations averaged remains the same.)  This agrees with the hypothesis for the flat sky case, for which e.g. Zaldarriaga (2001) \cite{Zaldarriaga2001a} argued that form of the radial function has to be ${}_\mm f(\beta>0)\simeq\beta^{-2}$ to ensure that the Fourier modes of Stokes $Q/U$ relate to those of $E/B$ merely by rotations (with no scale dependence). In addition, we find that ${}_{\md}f$ under a moving average shows a similar scaling of $\beta^{-2}$, as in \fig{fig:dfbeta_avg}.}

\revisit{In essence the radial function resulting from taking a sliding average can be thought of as transitioning from a higher band limit radial function to a radial function with an effectively lower band limit, set by the width of the sliding window.}
\comment{K: I don't follow this sentence.} 
\revisit{Applying a moving average on radial functions with increasing band limit, we see that a larger fraction of angular domain matches the scaling of $\beta^{-2}$.   Extrapolating this trend to the case of a very high band limit we may expect that this behavior will approach $\beta^{-2}$ across the domain $\beta\in[\epsilon,\pi-\epsilon]$ for $\epsilon \ll 1$.}

\revisit{Near the edges of the domain $\beta \rightarrow 0,\pi$ the window function partly falls outside the domain and this is in part responsible for the scaling deviating from the power law, but it is also important to note that ${}_{\mm}f$ and ${}_{\md}f$ need to vanish at $\beta=0$ and therefore the sliding average in the vicinity of these points cannot have the same power law behavior.}
